Predicting Post-Election FAANG Stock Prices: Accurate November Trends and Flat December Projections

### 2024-12-04 paper/paper.qmd
committing updated paper.qmd and paper.pdf

library(tidyverse)
library(ggplot2)
library(arrow)
library(kableExtra)
library(xgboost)
library(tidyr)
library(zoo)
library(scales)
library(reshape2)

@@

This paper leverages historical stock price data sourced from Yahoo Finance (2020–2024) to develop predictive models for FAANG stock prices in November and December 2024. Utilizing XGBoost regression models, we analyze historical trends and engineered features, including lagged prices, moving averages, and volatility measures, to forecast stock price movements. These predictions provide valuable insights into the behavior of FAANG stocks during critical periods, such as post-election market adjustments, and highlight opportunities and risks for short-term and long-term investment strategies.

@@

This analysis uses historical stock price data for FAANG companies (Meta, Amazon, Apple, Netflix, and Google) obtained from the Yahoo Finance API via the `tidyquant` library in R on November 30, 2024. The dataset spans from January 1, 2020, to December 31, 2024, providing daily records of stock performance, including key metrics such as opening, high, low, and closing prices, trading volume, and adjusted closing prices normalized for corporate actions. To enhance the predictive power of the data, several derived features were constructed. These include lagged adjusted prices to capture sequential trends, daily returns to measure proportional changes, rolling averages to smooth short- and long-term fluctuations, and a rolling standard deviation of prices to quantify market volatility. These features offer a structured framework for identifying patterns and forecasting future stock movements.

The raw data underwent pre-processing to address missing values, particularly for adjusted prices. Missing entries were filled using forward and backward imputation to ensure data continuity. Placeholder dates for December 2024 were interpolated to prepare for out-of-sample predictions. The cleaned data, prepared using the `tidyverse` of [@tidyverse], `lubridate` of [@lubridate], `TTR` of [@ttr], and `arrow` of [@arrow] libraries, provides a consistent and comprehensive foundation for modeling and analysis. Further details on the data preparation process are included in [Appendix -@sec-data-details].

## Measurement
Stock price data encapsulates market dynamics influenced by diverse factors, including macroeconomic conditions, corporate events, and investor sentiment. The raw variables in this dataset, such as **`open`**, **`high`**, **`low`**, **`close`**, and **`adjusted`**, capture key price points and provide a snapshot of market activity for each trading day. However, while these variables accurately reflect daily trading outcomes, they may not fully represent intraday fluctuations or complex trading dynamics.

Constructed variables enhance the analytical depth of the dataset by quantifying sequential trends, price variability, and momentum. For instance, **`Lag_1`** captures the adjusted closing price from the previous trading day, providing context for sequential patterns. Variables like **`daily_return`** measure proportional daily changes in price, while moving averages such as **`Rolling_Mean_7`**, **`sma_20`**, and **`sma_50`** smooth short- and long-term price fluctuations, offering a clearer view of overall trends. Additionally, **`volatility`**, computed as the rolling standard deviation of closing prices, quantifies recent price variability and reflects market uncertainty.

Market activity is further contextualized through **`volume`**, representing the total shares traded daily, a proxy for investor interest and liquidity. The inclusion of **`symbol_encoded`**, a numerical encoding of stock tickers, standardizes categorical data for compatibility with machine learning models. Together, these features create a robust framework for analysis, balancing raw price data with derived metrics that highlight market trends and dynamics.

Despite the strengths of these variables, limitations exist. Raw price points do not account for intraday variations, while constructed metrics may oversimplify complex market behaviors. External shocks, such as economic crises or regulatory changes, may not be fully captured within the variables’ scope. Acknowledging these constraints is essential for interpreting the results and understanding the broader context of the analysis. These limitations are further discussed in [Appendix -@sec-data-details].

@@

The dataset used for this analysis includes both original and constructed variables, designed to capture temporal trends, volatility, and momentum in stock price movements. Below is a detailed overview of the variables included in the model:

### Original Variables
The original variables directly obtained from the raw stock price data are:

@@

### Constructed Variables
The constructed variables were engineered to enhance the model's predictive performance by capturing price trends, volatility, and momentum:

@@

# Plot the distribution of the adjusted closing price, faceted by stock
ggplot(analysis_data, aes(x = adjusted)) +
  geom_histogram(bins = 30, fill = "blue", color = "black", alpha = 0.7) +
  facet_wrap(~symbol, scales = "free") +
  labs(
    x = "Adjusted Closing Price",
    y = "Frequency"
  ) +
  theme_minimal() +
  theme(
    text = element_text(size = 12),
    plot.title = element_text(hjust = 0.5, face = "bold"),
    strip.text = element_text(face = "bold")
  )
```
@fig-outcome-var illustrates the distribution of the adjusted closing prices (adjusted) for each FAANG stock, faceted by stock symbol. These distributions highlight the variability in stock prices across the companies, with stocks like NFLX showing higher price ranges, whereas GOOGL and AMZN have narrower distributions. The visual emphasizes the diverse price behavior among FAANG stocks, underscoring the need for tailored models that capture these unique trends. By normalizing for corporate actions like stock splits or dividends, the adjusted closing price serves as a reliable target variable for forecasting and analysis.

@@

Together, these variables provide a robust framework for understanding and predicting stock price movements over time, leveraging both raw market indicators and derived metrics.

@@

# Select and display random sample of predictive variables from analysis_data
set.seed(123) # For reproducibility
library(dplyr)
library(knitr)

@@

```
@tbl-pred-var displays a random selection of five rows from the dataset, showcasing the predictive variables utilized in the modeling process. Each row represents data for a specific stock, including values for lagged adjusted closing prices (Lag_1), smoothed weekly averages (Rolling Mean (7-Day)), and medium- and long-term moving averages (SMA (20-Day) and SMA (50-Day)). These variables are complemented by measures of price volatility, daily percentage returns (Daily Return), and an encoded numerical identifier for the stock symbol (Symbol (Encoded)). All values are rounded to two decimal places to enhance clarity and precision. This table illustrates the diverse set of historical and technical indicators used to predict future adjusted stock prices.

@@

The models in this study forecast the adjusted closing prices of FAANG stocks (Meta, Amazon, Apple, Netflix, and Google) for the months of November and December 2024. Separate models are constructed for each month to capture the time-dependent nature of stock price movements and to account for fluctuations in market conditions leading up to the forecast period. These models utilize a combination of historical stock price data and technical indicators, such as moving averages, volatility, and lagged price features, to estimate future stock prices for each of the five FAANG companies.

To model the changes in stock prices over time, we use features that capture both short-term and long-term market trends. The models incorporate lagged stock prices, 7-day rolling averages, 20-day and 50-day simple moving averages (SMA), daily returns, and volatility indicators. These features are crucial for understanding the patterns and cycles in stock price movements. XGBoost, a powerful machine learning algorithm, is used to build these models, allowing for the capture of non-linear relationships and interactions between the technical indicators and future stock prices.

The training datasets for the November and December models are defined using data up to October 31, 2024, and November 30, 2024, respectively. The models take into account temporal changes in stock prices by using features like lagged prices and rolling averages, which help capture the dynamics of stock price behavior. Additionally, XGBoost's flexibility allows the models to account for varying levels of volatility and market shifts across the different FAANG stocks.

This modeling approach enables the prediction of future stock prices, considering how past price trends and market indicators evolve. By focusing on the November and December timeframes, the models aim to provide forecasts that reflect recent market conditions and predict likely price movements as we approach the end of 2024. The detailed model specifications, including feature selection and hyperparameter tuning, are further discussed in [Appendix -@sec-model-details]

@@

To ensure the validity and reliability of our stock price prediction models, several key assumptions are made during the modeling process. These assumptions are necessary to capture the complexities of stock price behavior while maintaining a clear and interpretable framework for forecasting:

@@

These assumptions form the foundation of the models and guide the interpretation of the results. While they allow for effective forecasting, it is important to acknowledge that stock price prediction is inherently uncertain, and the accuracy of predictions may vary depending on unforeseen market events and fluctuations.

@@

The primary goal of the November model is to predict the adjusted closing prices of FAANG stocks for the month of November 2024 based on historical stock price data. This model is structured as a regression task using the XGBoost algorithm to estimate the future adjusted closing prices for each of the five FAANG companies.

The model setup can be described as follows:

@@

The dataset for the November model is prepared by extracting the relevant data from January 2020 to October 31, 2024. Features like moving averages and volatility are calculated based on this historical data, and the model is trained to predict the adjusted closing prices for November 2024.

The model was implemented using the **`xgboost`** package in R, which is optimized for performance in regression tasks:

$$
y_i = \text{XGBoost}(X_i, \theta)
$$

Where $y_i$ represents the actual adjusted closing price for day $i$ and $\theta$ represents the model parameters to be learned during training. The model is trained with the following key parameters:
- **Max Depth**: 6
- **Learning Rate**: 0.1
- **Number of Rounds**: 100
- **Objective Function**: Regression with squared error

@@

The December model follows a similar approach to the November model, but is trained using data up to November 30, 2024. This ensures that the model accounts for the most recent stock price movements leading up to December, providing a more accurate prediction for the final month of 2024.

The setup for the December model mirrors that of the November model, with the same features and target variable, but applied to the updated dataset. The model structure remains the same, and the XGBoost algorithm is again used for regression:

@@

The December model is trained using the same XGBoost parameters as the November model. Once trained, the model will generate predictions for December 2024, leveraging the latest market conditions reflected in the training data.

@@

The XGBoost model is the most suitable approach for forecasting FAANG stock prices in November and December 2024. Our primary outcome variable is the adjusted closing price of each stock, which is continuous and unbounded. XGBoost, a gradient boosting algorithm, excels in handling regression tasks with large datasets, capturing complex, non-linear relationships between features such as lagged prices, moving averages, volatility, and daily returns. To model these relationships effectively, XGBoost's ability to handle feature interactions and non-linearities is key. The inclusion of rolling averages and volatility allows the model to account for both short-term fluctuations and longer-term trends in stock prices, which are often influenced by broader market dynamics. Additionally, the model's flexibility helps capture the intricacies of stock price behavior over time, ensuring that past market patterns are reflected in future predictions.

@@

@@

Using the models trained on historical data, we applied the XGBoost algorithm to predict the adjusted closing prices for each of the FAANG stocks during these two months. For November, we filtered the data from November 1st to 30th, 2024, and prepared the relevant features for prediction, including lagged prices, rolling means, and volatility indicators. Since we have the actual adjusted closing prices for November, we compare the predicted values with the actual data to evaluate model performance. For December, the same approach was applied to the data from December 1st to 31st, 2024, but since the actual prices for this month are unavailable, the results for December focus solely on the predicted values.

@@

## Actual vs Predicted for November 2024
```{r}
#| echo: false
#| eval: true
#| label: fig-nov-actual-pred
#| fig-cap: "Actual vs Predicted Stock Prices for November 2024"
#| warning: false

# Plot actual vs predicted prices for November 2024
ggplot(nov_results, aes(x = as.Date(date))) +
  geom_line(aes(y = adjusted, color = "Actual"), size = 1) +  # Actual prices
  geom_line(aes(y = predicted_adjusted, color = "Predicted"), linetype = "dashed", size = 1) +  # Predicted prices
  facet_wrap(~symbol, scales = "free_y") +  # Separate panels for each stock
  labs(
    x = "Date",
    y = "Stock Price",
    color = "Legend"
  ) +
  scale_color_manual(values = c("Actual" = "blue", "Predicted" = "red")) +  # Define colors
  theme_minimal() +
  theme(
    legend.position = "bottom",
    text = element_text(size = 12),
    axis.text.x = element_text(angle = 45, hjust = 1)
  )
```
@fig-nov-actual-pred presents the comparison of actual versus predicted stock prices for the FAANG companies during November 2024. For each company—Apple (AAPL), Amazon (AMZN), Meta (META), Netflix (NFLX), and Google (GOOGL)—we plot the actual adjusted closing prices (in blue) alongside the predicted values (in red). The predictions, derived from our XGBoost models, demonstrate a generally strong alignment with the observed stock price movements, particularly for AAPL and META, where the predicted prices closely follow the actual trends. However, for companies like NFLX, the predicted values show a less consistent match, especially later in the month. This discrepancy suggests that while the model captures major trends effectively, there may be some volatility or market events that were not fully accounted for in the prediction process. These visualizations help assess the performance of our models and highlight areas where the predictions are robust as well as where there is room for improvement.

## Heatmap of Prediction Errors (November 2024)
```{r}
#| echo: false
#| eval: true
#| label: fig-heatmap
#| fig-cap: "Heatmap of Prediction Errors for November 2024"
#| warning: false

# Add the `error` column to `nov_results`
nov_results <- nov_results %>%
  mutate(error = abs(adjusted - predicted_adjusted))  # Calculate absolute error

# Prepare data for heatmap
nov_heatmap_data <- nov_results %>%
  select(date, symbol, error) %>%
  spread(symbol, error)

# Plot heatmap of prediction errors
ggplot(melt(nov_heatmap_data, id.vars = "date"), aes(x = date, y = variable, fill = value)) +
  geom_tile() +
  scale_fill_gradient2(low = "blue", mid = "white", high = "red", midpoint = 0) +
  labs(
    x = "Date",
    y = "Stock Symbol",
    fill = "Error"
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    text = element_text(size = 12)
  )
```
@fig-heatmap highlights the discrepancies between the predicted and actual stock prices for each FAANG company on a daily basis. The color intensity, ranging from light pink to deep red, indicates the magnitude of the errors, with darker shades representing higher errors. Notably, the Netflix (NFLX) stock shows consistently higher errors throughout November, particularly on November 4th and November 11th, where the error values reach above 100, as indicated by the deeper red color. Meta (META) and Google (GOOGL) exhibit moderate errors, with values peaking around November 11th but staying relatively consistent compared to NFLX. In contrast, Amazon (AMZN) and Apple (AAPL) show smaller error values, especially in the first half of the month, where the colors are much lighter. This heatmap provides a clear view of which stocks and dates the model struggled with the most, offering insights into potential areas for further model improvement.

## Average Predicted Stock Price - November
```{r}
#| echo: false
#| eval: true
#| label: fig-nov-avg
#| fig-cap: "Average Predicted Stock Prices for November 2024"
#| warning: false

@@

```
@fig-nov-avg illustrates the average predicted stock prices for each FAANG company in November 2024. The predictions reveal significant variations across the companies, with Netflix (NFLX) showing the highest predicted average price, approximately **$600**, followed by Meta (**'META'**) with an average just above **$500**. In contrast, Apple (**'AAPL'**), Amazon (**'AMZN'**), and Google (**'GOOGL'**) exhibit considerably lower average predicted prices, with AAPL at around **$225**, AMZN at **$210**, and GOOGL at approximately **$180**. This visualization highlights the distinct predicted performance of each company, suggesting that Netflix is expected to outperform the others in terms of stock price in November 2024, based on our model’s predictions. These results provide a clear comparative overview of expected stock movements and serve as a foundation for further analysis of market trends in the upcoming months.

@@

The line chart above presents the predicted stock prices for each FAANG company throughout December 2024. The predicted prices for each stock exhibit a strikingly consistent slope, with no significant fluctuations or adjustments throughout the month. For example, Netflix (NFLX) shows a steady increase, while Apple (AAPL), Amazon (AMZN), and Google (GOOGL) display relatively flat predicted price trends. This consistent pattern suggests potential limitations in the model, as real-world stock prices are typically subject to more volatility and market dynamics that are not captured by this prediction.

```{r}
#| echo: false
#| eval: true
#| label: fig-dec-pred
#| fig-cap: "Predicted Stock Price Trends for December 2024"
#| warning: false

ggplot(dec_results, aes(x = as.Date(date), y = predicted_adjusted, color = symbol)) +
  geom_line(size = 1) +
  labs(
    x = "Date",
    y = "Predicted Stock Price",
    color = "Stock Symbol"
  ) +
  theme_minimal() +
  theme(
    legend.position = "bottom",
    text = element_text(size = 12),
    axis.text.x = element_text(angle = 45, hjust = 1)
  )
```
The lack of variability in the predictions may indicate that the model fails to account for potential market shifts, such as reactions to economic news, company announcements, or other external factors. Additionally, the absence of any sharp price movements could suggest that the model overly smooths the data, possibly due to the features used or the nature of the training data. These limitations highlight the importance of refining the model to incorporate broader market influences and improve its responsiveness to sudden changes in the stock market environment.

@@

```{r}
#| echo: false
#| eval: true
#| label: fig-dec-avg
#| fig-cap: "Average Predicted Stock Prices for December 2024"
#| warning: false

@@

```
The bar chart above displays the average predicted stock prices for each FAANG company in December 2024. As shown, Netflix (NFLX) is expected to have the highest average predicted stock price, just above **$750**, followed by Meta (META) at approximately **$600**. In contrast, Apple (AAPL), Amazon (AMZN), and Google (GOOGL) show significantly lower predicted prices, with AAPL around **$250**, AMZN slightly lower, and GOOGL at just above **$200**. This suggests that, according to the model, Netflix is projected to continue outperforming other FAANG companies in terms of stock price. These predictions reflect market trends modeled from historical data, with a noticeable discrepancy between the higher-performing stocks and the others. The chart provides a comparative view of how each stock is expected to fare during December 2024, contributing to our understanding of potential market movements for the month.

## Average Predicted Stock Price - December

```{r}
#| echo: false
#| eval: true
#| label: tbl-dec-avg
#| tbl-cap: "Average Predicted Stock Prices for December 2024"
#| warning: false

# Calculate average predicted price for December
dec_avg_prices <- dec_results %>%
  group_by(symbol) %>%
  summarize(avg_predicted_price = mean(predicted_adjusted, na.rm = TRUE)) %>%
  mutate(avg_predicted_price = dollar(round(avg_predicted_price, 2))) %>%  # Format as dollar and round
  arrange(desc(avg_predicted_price))  # Sort by average price in descending order

# Display the table
dec_avg_prices %>%
  kable(
    caption = "Average Predicted Stock Prices for December 2024",
    col.names = c("Stock Symbol", "Average Predicted Price"),
    format = "markdown"  # Output as markdown for readability
  )
```
@tbl-dec-avg presents the average predicted stock prices for each FAANG company in December 2024. According to the predictions, Netflix (NFLX) is expected to have the highest average stock price at **$887.84**, followed by Meta (META) with an average of **$574.92**. Apple (AAPL) is predicted to have an average stock price of **$235.37**, while Amazon (AMZN) and Google (GOOGL) are expected to have lower average prices, at **$207.93** and **$168.69**, respectively. These predicted values highlight significant variations across the FAANG companies, with Netflix forecasted to substantially outperform the others. This table offers a clear numerical summary of the expected market performance for each company, providing insights into potential market trends for December 2024.

## Combined November and December
```{r}
#| echo: false
#| eval: true
#| label: fig-nov-dec-pred
#| fig-cap: "Predicted Stock Prices for November and December 2024"
#| warning: false

@@

```
@fig-nov-dec-pred above illustrates the predicted stock prices for each FAANG company across November and December 2024. The predictions are shown for both months, with the November predictions represented by the teal lines and the December predictions shown in red. For companies like Apple (AAPL) and Meta (META), the model predicts a notable increase in stock prices during November, followed by a flat trend throughout December, indicating minimal movement in the predicted prices during the second month. In contrast, for Netflix (NFLX), the predicted stock price remains almost flat through both months, showing limited variation. Amazon (AMZN) and Google (GOOGL) exhibit more fluctuations in November, with predicted prices stabilizing into flat trends for December. This pattern highlights a potential limitation of the model, as it fails to capture significant market dynamics or volatility in December, suggesting that the model may not fully account for market influences or external factors that could impact stock prices at the end of the year.

@@

```{r}
#| echo: false
#| eval: true
#| label: fig-nov-dec-avg
#| fig-cap: "Average Predicted Stock Prices for November and December 2024"
#| warning: false

@@

# Additional data details {#sec-data-details}

# Model details {#sec-model-details}

## Posterior predictive check

## Diagnostics

>>>>

library(tidyverse)
library(ggplot2)
library(arrow)
library(kableExtra)
library(xgboost)
library(tidyr)
library(zoo)
library(scales)
library(reshape2)
library(yardstick)

@@

This paper focuses on historical stock price data from years 2020 to 2024, sourced from Yahoo Finance [@yahoofinance] to develop predictive models for FAANG stock prices in November and December 2024. Utilizing XGBoost regression models, we analyze historical trends and engineered features, including lagged prices, moving averages, and volatility measures, to forecast stock price movements. These predictions provide valuable insights into the behavior of FAANG stocks during critical periods, such as post-election market adjustments, and highlight opportunities and risks for short-term and long-term investment strategies.

@@

This analysis uses historical stock price data for FAANG companies (Meta, Amazon, Apple, Netflix, and Google) obtained from the Yahoo Finance API [@yahoofinance] via the `tidyquant`of @tidyquant library in R [@citeR] on November 30, 2024. The data ranges from January 1, 2020, to November 29th, 2024, providing daily records of stock performance, including key metrics such as opening, high, low, and closing prices, trading volume, and adjusted closing prices normalized for corporate actions. To extend the data for December 2024, placeholder dates are generated and combined with the existing data, followed by interpolation to fill missing values in the adjusted price column. Such data cleaning ensures a complete and continuous dataset for model training and prediction. Details of the data cleaning and interpolation process are provided in [Appendix -@sec-data-cleaning].

Data cleaning and analysis are conducted by employing the following packages: 'tidyverse' of @tidyverse, 'TTR' of @ttr, and 'arrow' of @arrow.

## Measurement

Stock prices reflect a company's perceived value in financial markets, influenced by factors like economic conditions, corporate performance, and investor sentiment. These dynamics are captured through daily trading data, where variables such as **'open'**, **'high'**, **'low'**, **'close'**, **'adjusted'**, and **'volume'** provide a structured snapshot of market activity. Each data point represents an aggregated view of trading outcomes for a given day, simplifying continuous price fluctuations into discrete records. For example, the adjusted price accounts for corporate actions like stock splits and dividends, offering a standardized metric for historical comparison. Volume indicates total shares traded, reflecting market liquidity and investor interest but not the motivations behind trades.

The data assumes consistency across trading days, where daily aggregates like close encapsulate the final market consensus for that day. However, this measurement approach may overlook intraday volatility and the nuanced impact of macroeconomic events or breaking news that influence trading behavior. Similarly, variables like date simplify time into discrete intervals, capturing trends over time but not the continuous nature of market dynamics. While these measurements are essential for analyzing historical trends, they simplify complex phenomena like investor sentiment and market reactions, such as the rationale behind trades or varying confidence levels. To address these limitations, feature engineering constructs variables that capture trends, volatility, and other patterns, enhancing the data’s analytical value for model training and predictive analysis. Further details are provided in [Appendix -@sec-data-details].

@@

The collected data from Yahoo Finance [@yahoofinance] contains several key variables relevant to the analysis.

@@

### Constructed Variables
To enhance the data's analytical depth and support predictive modeling, the following constructed variables were engineered to capture price trends, volatility, and momentum:

@@

# Plot the distribution of the adjusted closing price, faceted by stock
ggplot(analysis_data, aes(x = adjusted)) +
  geom_histogram(bins = 30, fill = "blue", color = "black", alpha = 0.7) +
  facet_wrap(~symbol, scales = "free") +
  labs(
    x = "Adjusted Closing Price",
    y = "Frequency"
  ) +
  theme_minimal() +
  theme(
    text = element_text(size = 12),
    plot.title = element_text(hjust = 0.5, face = "bold"),
    strip.text = element_text(face = "bold")
  )
```
@fig-outcome-var illustrates the distribution of the adjusted closing prices (**'adjusted'**) for each FAANG stock, faceted by stock symbol. These distributions highlight the variability in stock prices across the companies, with stocks like NFLX showing higher price ranges, whereas GOOGL and AMZN have narrower distributions. The visual emphasizes the diverse price behavior among FAANG stocks, underscoring the need for tailored models that capture these unique trends. By normalizing for corporate actions like stock splits or dividends, the adjusted closing price serves as a reliable target variable for forecasting and analysis.

@@

Together, these variables provide a robust framework for understanding and predicting stock price movements over time, utilizing both raw market indicators and derived metrics.

@@

# Select and display random sample of predictive variables from analysis_data
set.seed(123) # For reproducibility

@@

```
@tbl-pred-var displays a random selection of five rows from the dataset, showcasing the predictive variables utilized in the modeling process. Each row represents data for a specific stock, including values for lagged adjusted closing prices (Lag_1), smoothed weekly averages (Rolling Mean (7-Day)), and medium- and long-term moving averages (SMA (20-Day) and SMA (50-Day)). These variables are complemented by measures of price volatility, daily percentage returns (Daily Return), and an encoded numerical identifier for the stock symbol (Symbol (Encoded)). All values are rounded to two decimal places to enhance clarity and precision.

@@

The XGBoost regression models in this study predict adjusted closing prices for FAANG stocks—Meta, Amazon, Apple, Netflix, and Google—for November and December 2024. Separate models are constructed for each month to account for time-specific market dynamics and recent trends. To capture short and long-term trends in stock price movements, these models incorporate features such as lagged prices, moving averages, daily returns, and volatility. Training data is segmented by timeframe, with the November model using data up to October 31, 2024, and the December model using data through November 30, 2024. Temporal features and technical indicators enable the models to capture stock-specific volatility and broader market trends. Further details on model specifications and diagnostics are provided in [Appendix -@sec-model-details].

@@

To ensure the validity and reliability of our XGBoost models, several key assumptions are made during the modeling process:

@@

@@

The primary estimand of the November model is the adjusted closing price of FAANG stocks for November 2024. The model is structured as a regression task using the XGBoost algorithm to estimate these prices for each of the five FAANG companies. The model setup is as follows:

@@

The data for the November model is prepared by extracting the relevant data from January 2020 to October 31, 2024. Features like moving averages and volatility are calculated based on this historical data, and the model is trained to predict the adjusted closing prices for November 2024.

The model was implemented using the `xgboost` of @xgboost in R, which is optimized for performance in regression tasks:

$$
y_i = \text{XGBoost}(X_i, \theta)
$$
Where $y_i$ represents the actual adjusted closing price for day $i$ and $\theta$ represents the model parameters to be learned during training.

Key hyperparameters for the model are specificed as:

- **Max Depth**: 6, to control the complexity of the trees and prevent overfitting
- **Learning Rate**: 0.1, to balance the contribution of each boosting iteration
- **Number of Rounds**: 100, to determine the total iterations for model optimization
- **Objective Function**: Regression with squared error, suitable for continuous target variables.

This set-up ensures that the model captures patterns in stock price movements while balancing bias and variance for improved predictive performance. Key assumptions include the appropriateness of historical stock price trends and technical indicators for predicting future prices.

@@

The December model follows a similar approach to the November model but incorporates extrapolated data for December 2024. Placeholder dates for December are added, and missing values in the adjusted price column are filled using forward and backward interpolation. This ensures the model reflects recent trends while accounting for the data cutoff on November 30, 2024. The model structure remains the same, and the XGBoost algorithm is again used for regression:

@@

The December model is trained using the `XGBoost` of @xgboost library and the same parameters as the November model. Once trained, the model will generate predictions for December 2024, leveraging the latest market conditions reflected in the training data.

@@

The XGBoost model is the most suitable approach for forecasting FAANG stock prices in November and December 2024. Our primary outcome variable is the adjusted closing price of each stock, which is continuous and unbounded. XGBoost, a gradient boosting algorithm, excels in handling regression tasks with large datasets, capturing complex, non-linear relationships between features such as lagged prices, moving averages, volatility, and daily returns. To model these relationships effectively, XGBoost's ability to handle feature interactions and non-linearities is key. The inclusion of rolling averages and volatility allows the model to account for both short-term fluctuations and longer-term trends in stock prices, which are often influenced by broader market dynamics. Additionally, the model's flexibility helps capture the intricacies of stock price behavior over time, ensuring that past market patterns are reflected in future predictions. Details are further discussed in [Appendix -@sec-model-justification]

@@

## Model summary
Both models were evaluated using RMSE, MAE, and R² metrics, as shown below in @fig-model-summary. The November model achieved an RMSE of 40.50, MAE of 16.40, and R² of 0.988, indicating it captured meaningful variability in the stock price movements. The December model reported significantly lower RMSE (1.10) and MAE (0.78), with an R² of 0.999, suggesting near-perfect alignment with the data. However, the December model's near-flat projections likely stem from minimal variability in the placeholder data used for December, resulting in artificially inflated performance metrics.
```{r}
#| echo: false
#| eval: true
#| label: fig-model-summary
#| fig-cap: "November and December Model Summary"
#| warning: false

# Filter test data for November and December
nov_test <- analysis_data %>%
  filter(as.Date(date) > as.Date("2024-10-31") & as.Date(date) <= as.Date("2024-11-30"))

dec_test <- analysis_data %>%
  filter(as.Date(date) > as.Date("2024-11-30") & as.Date(date) <= as.Date("2024-12-31"))

# Prepare test features and labels for November
X_test_nov <- nov_test %>%
  select(Lag_1, Rolling_Mean_7, sma_20, sma_50, volatility, daily_return, symbol_encoded) %>%
  as.matrix()
y_test_nov <- as.numeric(nov_test$adjusted)

# Prepare test features and labels for December
X_test_dec <- dec_test %>%
  select(Lag_1, Rolling_Mean_7, sma_20, sma_50, volatility, daily_return, symbol_encoded) %>%
  as.matrix()
y_test_dec <- as.numeric(dec_test$adjusted)

# Generate predictions
nov_preds <- predict(nov_model, X_test_nov)
dec_preds <- predict(dec_model, X_test_dec)

# Calculate metrics for November
nov_metrics <- tibble(
  Metric = c("RMSE", "MAE", "R²"),
  November = c(
    rmse_vec(y_test_nov, nov_preds),
    mae_vec(y_test_nov, nov_preds),
    rsq_vec(y_test_nov, nov_preds)
  )
)

# Calculate metrics for December
dec_metrics <- tibble(
  Metric = c("RMSE", "MAE", "R²"),
  December = c(
    rmse_vec(y_test_dec, dec_preds),
    mae_vec(y_test_dec, dec_preds),
    rsq_vec(y_test_dec, dec_preds)
  )
)

# Combine metrics into one table
combined_metrics <- nov_metrics %>%
  left_join(dec_metrics, by = "Metric")

# Display combined table using kable
kable(combined_metrics)
```
These results highlight the November model's relative robustness in capturing dynamic trends compared to the December model's overfitted nature due to static input data. This emphasizes the importance of data variability in evaluating predictive model performance.

@@

Using the models trained on historical data, we implemented the XGBoost algorithm to predict the adjusted closing prices for each of the FAANG stocks during these two months. For November, since we have the actual adjusted closing prices for November, we compare the predicted values with the actual data to evaluate model performance. For December, since the actual prices for this month are unavailable, the results for December focus solely on the predicted values.

@@

## Actual vs Predicted for November 2024
```{r}
#| echo: false
#| eval: true
#| label: fig-nov-actual-pred
#| fig-cap: "Comparison of actual (blue) vs. predicted (red) adjusted closing stock prices for FAANG companies in November 2024. Highlights the model's ability to capture general trends but shows variability in performance across stocks, particularly for Netflix (NFLX)"
#| warning: false

# Plot actual vs predicted prices for November 2024
ggplot(nov_results, aes(x = as.Date(date))) +
  geom_line(aes(y = adjusted, color = "Actual"), size = 1) +  # Actual prices
  geom_line(aes(y = predicted_adjusted, color = "Predicted"), linetype = "dashed", size = 1) +  # Predicted prices
  facet_wrap(~symbol, scales = "free_y") +  # Separate panels for each stock
  labs(
    x = "Date",
    y = "Stock Price",
    color = "Legend"
  ) +
  scale_color_manual(values = c("Actual" = "blue", "Predicted" = "red")) +  # Define colors
  theme_minimal() +
  theme(
    legend.position = "bottom",
    text = element_text(size = 12),
    axis.text.x = element_text(angle = 45, hjust = 1)
  )
```
@fig-nov-actual-pred compares actual (blue) and predicted (red) stock prices for FAANG companies during November 2024. The model captures trends effectively for AAPL and META, where predictions align closely with observed prices. However, for NFLX, discrepancies appear later in the month, indicating the model's limitations in capturing certain market volatilities.

## Average Predicted Stock Price - November
```{r}
#| echo: false
#| eval: true
#| label: fig-nov-avg
#| fig-cap: "Average predicted stock prices for FAANG companies in November 2024. Netflix leads with the highest predicted average price (~$750), while Google has the lowest (~$180), reflecting varied performance across companies based on model predictions."
#| warning: false

@@

```
@fig-nov-avg shows the average predicted stock prices for FAANG companies in November 2024. Netflix (NFLX) has the highest predicted average price (~$750), followed by Meta (META) at just above $500. Apple (AAPL), Amazon (AMZN), and Google (GOOGL) exhibit lower averages, at ~$225, ~$210, and ~$180, respectively. The chart highlights Netflix's expected strong performance compared to other FAANG companies.

@@

```{r}
#| echo: false
#| eval: true
#| label: fig-dec-pred
#| fig-cap: "Predicted stock prices for FAANG companies in December 2024. The model predicts relatively flat price trends for all companies, with Netflix having the highest predicted price and Google the lowest"
#| warning: false

ggplot(dec_results, aes(x = as.Date(date), y = predicted_adjusted, color = symbol)) +
  geom_line(size = 1) +
  labs(
    x = "Date",
    y = "Predicted Stock Price",
    color = "Stock Symbol"
  ) +
  theme_minimal() +
  theme(
    legend.position = "bottom",
    text = element_text(size = 12),
    axis.text.x = element_text(angle = 45, hjust = 1)
  )
```
@fig-dec-pred presents the predicted stock prices for FAANG companies in December 2024, showing flat trends with no significant fluctuations. Netflix (NFLX) has the highest predicted price, while Google (GOOGL) has the lowest. Such flat trends highlight potential limitations in the model's ability to capture real-world stock price volatility, further detailed in [Appendix -@sec-model-limitations].

@@

```{r}
#| echo: false
#| eval: true
#| label: fig-dec-avg
#| fig-cap: "Average predicted stock prices for FAANG companies in December 2024. Netflix leads with the highest predicted price ($825), followed by Meta ($600), while Apple, Amazon, and Google show considerably lower averages."
#| warning: false

@@

```
@fig-dec-avg shows the average predicted December 2024 stock prices for FAANG companies. Netflix (NFLX) leads with the highest predicted price, just around $825, followed by Meta (META) at $600. Apple (AAPL), Amazon (AMZN), and Google (GOOGL) are projected significantly lower, around $250, $200, and $200, respectively. These predictions highlight Netflix's strong expected performance compared to its peers, reflecting trends modeled from historical data.

## Combined November and December
```{r}
#| echo: false
#| eval: true
#| label: fig-nov-dec-pred
#| fig-cap: "Predicted stock prices for FAANG companies across November (red) and December (teal) 2024. November predictions exhibit variability, particularly for Apple (AAPL), Meta (META), Amazon (AMZN), and Google (GOOGL), while December predictions flatten, showing minimal fluctuations. Netflix (NFLX) remains nearly flat across both months. Highlights the model's reduced sensitivity to potential market dynamics in December."
#| warning: false

@@

```
@fig-nov-dec-pred illustrates predicted FAANG stock prices for November (teal) and December (red) 2024. November shows dynamic trends for Apple (AAPL), Meta (META), Amazon (AMZN), and Google (GOOGL), while December predictions flatten across all companies, particularly Netflix (NFLX). This highlights the model's limitation in capturing market volatility and end-of-year dynamics.

@@

```{r}
#| echo: false
#| eval: true
#| label: fig-nov-dec-avg
#| fig-cap: "Average predicted stock prices for FAANG companies in November (red) and December (teal) 2024. Netflix shows the highest increase between months, while Apple, Amazon, and Google exhibit stable predicted prices. Only google displays month-to-month decrease."
#| warning: false

@@

# Additional Data Details {#sec-data-details}

## Data Cleaning {#sec-data-cleaning}
